*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-08 Code/Manuscript_O2.Rmd
working on FtsH terminology

When assessing the presence of a particular protein complex at a sampling location, all peptides belonging to all subunits of the complex were included to give the greatest number of data points. As this data was acquired on a discovery mission rather than through targeted peptide approaches, it is difficult to discern accuracies of strain assignment annotations, particularly as the proteins of interest in this study are highly conserved across strains. We are, however, confident in clade classifications for each protein examined. A caveat to interpretation of this data is the peptide detection bias inherent to mass spectrometry XXX Aurora CITATION XXX. The data is also limited by the number and nature of protein spectra in the SEQUEST database: a peptide sequence was not determined unless there is already a known spectrum for that peptide in the SEQUEST database, hence some peptides of interest may not be identifiable. Furthermore, a peptide must be detected above a certain threshold abundance in order to be considered an accurate ‘hit’.

@@

Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Photosystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from clade LLII/III (ex. SS120) and clade LLIV (ex. MIT9313) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. The exclusive assignment of ATP Synthase peptides to clade LLIV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

XXX More Text XXX from Amanda & Aurora XXX
Proteins derived from clade HLXXX ecotypes of *P. marinus* were detected in OMZ at depths up to 200 meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines the niches occupied by different *P. marinus* ecotypes, as compared to potentially covarying environmental variables like photoperiod, light spectrum, and light level, is poorly described.

@@

Under full atmospheric [O~2~] and blue light, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. Repair of photoinactivated PSII relies on the removal of damaged PsbA [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition [@kanervoD1ProteinDegradation1993], mediated largely by a heterohexamer of (FtsH1-FtsH2)~3~, a membrane-bound [@sacharzSubCellularLocation2015; @zakInitialStepsBiogenesis2001] metalloprotease  [(Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014),  @boehmSubunitOrganizationSynechocystis2012a;  @komendaExposedNTerminalTail2007],  XXX (AMANDA REF from ALGATECH) XXX.

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light, HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* *slr*1463, possibly involved in PSI biogenesis, and FtsH expression did not increase in response to light stress in MIT9313 [@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

>>>>

When assessing the presence of a particular protein complex at a sampling location, all peptides belonging to all subunits of the complex were included to give the greatest number of data points. As this data was acquired on a discovery mission rather than through targeted peptide approaches, it is difficult to discern accuracies of strain assignment annotations, particularly as the proteins of interest in this study are highly conserved across strains. We are, however, confident in clade classifications for each protein examined. The data is also limited because a peptide sequence was not determined unless there was already a known spectrum for that peptide in the SEQUEST database, hence some peptides of interest may not be identifiable. Furthermore, a peptide must be detected above a certain threshold abundance in order to be considered an accurate ‘hit’.

@@

Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on core photosynthetic protein complexes, for clades HL, LLI, LLIII and LLIV (Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (including strain MED4) were detected throughout the water column, predominantly at high [O~2~]. The  absence of proteins annotated for the key ATP Synthase complex for Clade HLI, compared to annotated detections of ATP Synthase across the other three clades suggests limitations in the annotation process for highly conserved protein sequences. Complexes from Clades LLI (including strain NATL2A) were also present across the depth/light axes, with more representation at lower [O~2], and fewer near surface detections. Complexes from clade LLII/III (ex. SS120) were also detected across the depth/light and [O~2~] ranges, with more detections at deeper, darker depths.  Clade LLIV (ex. MIT9313) photosynthetic complexes were also were detected throughout the depth/light and [O~2~], ranges, with the most frequent detections at depth and at low [O~2~], compared to other strains.

@@

Under full atmospheric [O~2~] and blue light, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. Repair of photoinactivated PSII relies on the removal of damaged PsbA [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition [@kanervoD1ProteinDegradation1993], mediated largely by a heterohexamer of (FtsH1-FtsH2)~3~, a membrane-bound [@sacharzSubCellularLocation2015; @zakInitialStepsBiogenesis2001] metalloprotease  [(Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014),  @boehmSubunitOrganizationSynechocystis2012a;  @komendaExposedNTerminalTail2007].

XXXStillWorkingonFtsH Terminology!XXXX
*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light, HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* *slr*1463, possibly involved in PSI biogenesis, and FtsH expression did not increase in response to light stress in MIT9313 [@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.
